---
title: "DATA624 Knowledge and Visual Analytics Module 1"
author: "Omar Pineda"
output:
  html_document: default
  html_notebook: default
---

**Principles of Data Visualization and Introduction to ggplot2**

I have provided you with data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. lets read this in:

```{r}

inc <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)

```

And lets preview this data:

```{r}

head(inc)

```
```{r}

summary(inc)

```

Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:

First, we load our libraries:

```{r load}
library(psych)
library(tidyverse)
```


The psych library helps us pull some additional exploratory information as well as statistics by each industry.

```{r}
describe(inc)
describeBy(inc, inc$Industry)
```

## Question 1

Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (ie taller than wide), which should further guide your layout choices.

```{r}
ggplot(inc, aes(inc$State)) +
  geom_bar(fill = "#0073C2FF", width = 1) +
  ylab("Number of companies") +
  xlab("State") +
  ggtitle("Number of companies in each U.S. state") +
  theme(text=element_text(size=6), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  coord_flip()
```

## Question 2

Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.

The state with the 3rd most companies in the data set is NY, which has 311 of the fastest growing companies.

```{r}
stateCount <- as.data.frame(table(inc$State))
head(stateCount[order(-stateCount$Freq),])
```

Next, we remove outliers and also calculate the average number of employees by industry for these NY companies to get an idea of what numbers we're working with. Energy and Financial Services have the highest average number of employees.

```{r}
#subsetting to NY companies and complete cases
ny <- inc[which(inc$State == 'NY'),]
ny <- ny[complete.cases(ny),]

#removing outliers as shown here https://stackoverflow.com/questions/53201016/remove-outliers-in-r-very-easy
b<-boxplot(ny$Employees)
lowerwhisker<-b$stats[1]
upperwhisker<-b$stats[5]
ny <- ny[ny$Employees>lowerwhisker & ny$Employees<upperwhisker,]

indavg <- ny %>%
    group_by(Industry) %>%
    dplyr::summarize(Mean = mean(Employees))
indavg
```

Lastly, we graph the distributions of employee counts for fast growth companies in New York by industry. It seems like Financial Services and Energy have, on average, more employees amongst NY's fast growth companies.

```{r}
ggplot(ny, aes(ny$Industry, ny$Employees)) +
  geom_boxplot() +
  ylab("Number of employees") +
  xlab("Industry") +
  ggtitle("Distribution of employees for fast growth companies in NY by industry") +
  theme(text=element_text(size=6), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  coord_flip()
```

```{r}
#ggplot(indavg, aes(indavg$Industry, indavg$Mean)) +
#  geom_bar(stat="identity", fill = "#0073C2FF", width = 1) +
#  ylab("Average number of employees") +
#  xlab("Industry") +
#  ggtitle("Average number of employees for fast growth companies in NY by industry") +
#  theme(text=element_text(size=6), panel.grid.major = element_blank(), panel.grid.minor = #element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
#  coord_flip()
```

## Question 3

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown.

First, we calculate average revenue by employee for each company, and then we calculate the average of this per capita value for each industry. This uses the dataset that had already removed outliers in the number of employees.

```{r}
ny$revenuebyemp <- ny$Revenue / ny$Employees

indAvgRev <- ny %>%
    group_by(Industry) %>%
    dplyr::summarize(Mean = mean(revenuebyemp))
indAvgRev
```

Finally, we visualize these figures in a chart. Energy generates the most revenue per employee compared to other industries for fast growth companies in NY.

```{r}
ggplot(ny, aes(ny$Industry, ny$revenuebyemp)) +
  geom_boxplot() +
  ylab("Average revenue per employee") +
  xlab("Industry") +
  ggtitle("Average revenue per employee for fast growth companies in NY by industry") +
  theme(text=element_text(size=6), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  coord_flip()
```
```{r}
#ggplot(indAvgRev, aes(indAvgRev$Industry, indAvgRev$Mean)) +
#  geom_bar(stat="identity", fill = "#0073C2FF", width = 1) +
#  ylab("Average revenue per employee") +
#  xlab("Industry") +
#  ggtitle("Average revenue per employee for fast growth companies in NY by industry") +
#  theme(text=element_text(size=6), panel.grid.major = element_blank(), panel.grid.minor = #element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) +
#  coord_flip()
```